### 2022-10-19 draft/draft.Rmd
almost done

@@

In this setting, if we know the labels $z_1, ..., z_n$, then for each $k$^th^ subsample for which $z_i = k$, we can fit the curve $\hat{g}_k$ (up to an indefinite orthogonal transformation). 
On the other hand, if we know the curves $g_1, ..., g_K$, then we can assign each embedding vector $\hat{x}_i$ the predicted label $\hat{z}_i$ based on which curve it is closest to. 
This lends itself to an alternating coordinate descent algorithm in which we alternate between fitting curves and assigning labels at each iteration (algorithm \ref{alg:kcurves}). 
We call this algorithm $K$-curves clustering. 
In this algorithm, the loss function that we wish to minimize is
\begin{equation}
\label{eq:kcurves-loss}
\sum_{k=1}^K \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2
\end{equation}

@@

\begin{theorem}
\label{thm:kcurves-semisupervised}
Let $A \sim \MBM(\{\alpha_1, ..., \alpha_K\}, \{F, ..., F\}, \{g_1, ..., g_K\}; \rho_n)$ such that $F$ has support $[0, 1]$ and its density function is such that $f(t) > a > 0$ for every $t \in [0, 1]$, and each $g_k$ is is a Bezier curve of order $R$. 
Suppose that for each community $k$, we have labels for $R + 1$ vertices. 
Define the community detection error rate as $\frac{1}{n} \min_\pi \sum_i I(z_i = \pi(\hat{z}_i))$, where $\pi$ is a permutation function and $\hat{z}_i$ are the predicted labels from $K$-curves clustering. 
Then as $n \to \infty$, $K$-curves clustering produces zero community detection error rate, with high probability. 
\end{theorem}

@@

In the final example,

>>>>

Manifolds have been observed in the ASEs of real observed graphs when they are studied as GRDPGs. 
The *Drosophila* connectome graph \citep{Eichler141762} displays community-wise linear and quadratic structures in the ASE. 
\citet{https://doi.org/10.48550/arxiv.1705.03297} used GMMs on the ASE for community detection and found that the curve-like structures in the ASE were incompatible with centroid-based clustering methods. 
\citet{SannaPassino2022} used Gaussian Processes to estimate curves, which resulted in better empirical fits. 
\citet{athreya2020estimation} considered GRDPGs in which the vertices lie on curves or manifolds, but their focus was largely on estimating distributions on the manifolds rather than the manifolds themselves.

Outside of random graph inference, there is a larger body of work on fitting parametric curves to vectors in $\mathbb{R}^d$ and identifying vectors belonging to different manifolds in $\mathbb{R}^d$.

@@

In this setting, if we know the labels $z_1, ..., z_n$, then for each $k$^th^ subsample for which $z_i = k$, we can fit the curve $\hat{g}_k$ (up to an indefinite orthogonal transformation). 
On the other hand, if we know the curves $g_1, ..., g_K$, then we can assign each embedding vector $\hat{x}_i$ the predicted label $\hat{z}_i$ based on which curve it is closest to. 
This lends itself to an alternating coordinate descent algorithm in which we alternate between fitting curves and assigning labels at each iteration (algorithm \ref{alg:kcurves}). 
We call this algorithm $K$-curves clustering. 
In this algorithm, the loss function that we wish to minimize is
\begin{equation}
\label{eq:kcurves-loss}
L(z_1, ..., z_n, g_1, ..., g_n; X) = \sum_{k=1}^K \sum_{i: z_i = k} \|x_i - g_k(t_i)\|^2
\end{equation}

@@

\begin{theorem}
\label{thm:kcurves-semisupervised}
Let $A \sim \MBM(\{\alpha_1, ..., \alpha_K\}, \{F, ..., F\}, \{g_1, ..., g_K\}; \rho_n)$ such that $F$ has support $[0, 1]$ and its density function is such that $f(t) > a > 0$ for every $t \in [0, 1]$, and each $g_k$ is is a Bezier curve of order $R$. 
Suppose that for each community $k$, we have labels for at least $R + 1$ vertices. 
Then as $n \to \infty$, the estimates outputted by $K$-curves clustering are such that 
$$L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_n; X) \stackrel{p}{\to} 0,$$
where $L$ is defined as in equation \ref{eq:kcurves-loss}. 
\end{theorem}

@@

In the final example, we examined the *Drosophila* connectome graph \citep{Eichler141762}, which has been studied as a GRDPG by \citet{athreya2017statistical}, \citet{https://doi.org/10.48550/arxiv.1705.03297}, and \citet{SannaPassino2022}. 
In this graph, each vertex is a neuron, and the labels correspond to one of four neuron types, Kenyon Cells, Input Neurons, Output Neurons, and Projection Neurons. 
In our analysis, we only considered neurons labeled as Kenyon Cells and Projection Neurons in the right hemisphere. 
The resulting graph is illustrated in figure \ref{fig:mbconnectome-graph}.

The ASE in two dimensions suggest that the MBM with one dimensional quadratic manifolds is an appropriate model for this graph (figure \ref{fig:mbconnectome-ase}). 
Fitting two quadratic Bezier curves to the ASE (figure \ref{fig:mbconnectome-kcurves}) results in 91.4\% accuracy.

```{r mbconnectome-graph, fig.cap = 'Drosophila connectome graph. Each vertex represents a neuron, which are labeled by neuron type.', fig.height = 2, figh.width = 4}
library(mbstructure)
data(MBconnectome)

graph <- generate.graph(newrdat, vdf.right)
A <- as.matrix(igraph::as_adjacency_matrix(graph$g, type = 'both'))
z <- graph$vdf$type
ind.keep <- z %in% c('KC', 'PN')
A <- A[ind.keep, ind.keep]
z <- as.numeric(as.factor(as.numeric(z[ind.keep])))
A <- sign(A + t(A))
diag(A) <- 0

```{r mbconnectome-ase, fig.cap = 'ASE of the Drosophila connectome graph. The embedding vectors are labeled by neuron type.', fig.height = 2, figh.width = 4}
Xhat <- embedding(A, 2, 0)

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
                 colour = factor(z))) + 
  coord_fixed() +
  theme_bw() + 
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = expression(x[1]), y = expression(x[2]),
       colour = NULL)
```

```{r mbconnectome-kcurves, fig.cap = 'ASE of the Drosophila connectome graph with quadratic Bezier curves fitted via $K$-curves clustering. The embedding vectors are colored by predicted label.', fig.height = 2, fig.width = 4, cache = TRUE}
set.seed(314159)
z.init <- sample.points(z, 3)

mbconnectome.out <- manifold.clustering(Xhat, 2, 
                              initialization = z.init,
                              intercept = FALSE, 
                              parallel = TRUE, 
                              verbose = FALSE, 
                              animate = FALSE, 
                              curve.init = 'x')

plot.estimated.curves(Xhat, mbconnectome.out) + 
  scale_colour_brewer(palette = 'Set1') + 
  # coord_fixed(.5) + 
  labs(x = expression(x[1]), y = expression(x[2]))
```

### 2022-10-19 draft/draft.tex
almost done

@@

In this setting, if we know the labels \(z_1, ..., z_n\), then for each \(k\)\textsuperscript{th} subsample for which \(z_i = k\), we can fit the curve \(\hat{g}_k\) (up to an indefinite orthogonal transformation).
On the other hand, if we know the curves \(g_1, ..., g_K\), then we can assign each embedding vector \(\hat{x}_i\) the predicted label \(\hat{z}_i\) based on which curve it is closest to.
This lends itself to an alternating coordinate descent algorithm in which we alternate between fitting curves and assigning labels at each iteration (algorithm \ref{alg:kcurves}).
We call this algorithm \(K\)-curves clustering.
In this algorithm, the loss function that we wish to minimize is
\begin{equation}
\label{eq:kcurves-loss}
\sum_{k=1}^K \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2
\end{equation}

@@

In the final example,

>>>>

Manifolds have been observed in the ASEs of real observed graphs when they are studied as GRDPGs.
The \emph{Drosophila} connectome graph \citep{Eichler141762} displays community-wise linear and quadratic structures in the ASE.
\citet{https://doi.org/10.48550/arxiv.1705.03297} used GMMs on the ASE for community detection and found that the curve-like structures in the ASE were incompatible with centroid-based clustering methods.
\citet{SannaPassino2022} used Gaussian Processes to estimate curves, which resulted in better empirical fits.
\citet{athreya2020estimation} considered GRDPGs in which the vertices lie on curves or manifolds, but their focus was largely on estimating distributions on the manifolds rather than the manifolds themselves.

Outside of random graph inference, there is a larger body of work on fitting parametric curves to vectors in \(\mathbb{R}^d\) and identifying vectors belonging to different manifolds in \(\mathbb{R}^d\).

@@

In this setting, if we know the labels \(z_1, ..., z_n\), then for each \(k\)\textsuperscript{th} subsample for which \(z_i = k\), we can fit the curve \(\hat{g}_k\) (up to an indefinite orthogonal transformation).
On the other hand, if we know the curves \(g_1, ..., g_K\), then we can assign each embedding vector \(\hat{x}_i\) the predicted label \(\hat{z}_i\) based on which curve it is closest to.
This lends itself to an alternating coordinate descent algorithm in which we alternate between fitting curves and assigning labels at each iteration (algorithm \ref{alg:kcurves}).
We call this algorithm \(K\)-curves clustering.
In this algorithm, the loss function that we wish to minimize is
\begin{equation}
\label{eq:kcurves-loss}
L(z_1, ..., z_n, g_1, ..., g_n; X) = \sum_{k=1}^K \sum_{i: z_i = k} \|x_i - g_k(t_i)\|^2
\end{equation}

@@

In the final example, we examined the \emph{Drosophila} connectome graph \citep{Eichler141762}, which has been studied as a GRDPG by \citet{athreya2017statistical}, \citet{https://doi.org/10.48550/arxiv.1705.03297}, and \citet{SannaPassino2022}.
In this graph, each vertex is a neuron, and the labels correspond to one of four neuron types, Kenyon Cells, Input Neurons, Output Neurons, and Projection Neurons.
In our analysis, we only considered neurons labeled as Kenyon Cells and Projection Neurons in the right hemisphere.
The resulting graph is illustrated in figure \ref{fig:mbconnectome-graph}.

The ASE in two dimensions suggest that the MBM with one dimensional quadratic manifolds is an appropriate model for this graph (figure \ref{fig:mbconnectome-ase}).
Fitting two quadratic Bezier curves to the ASE (figure \ref{fig:mbconnectome-kcurves}) results in 91.4\% accuracy.

{\centering \includegraphics{draft_files/figure-latex/mbconnectome-graph-1}

\caption{Drosophila connectome graph. Each vertex represents a neuron, which are labeled by neuron type.}\label{fig:mbconnectome-graph}
\end{figure}

{\centering \includegraphics{draft_files/figure-latex/mbconnectome-ase-1}

\caption{ASE of the Drosophila connectome graph. The embedding vectors are labeled by neuron type.}\label{fig:mbconnectome-ase}
\end{figure}

{\centering \includegraphics{draft_files/figure-latex/mbconnectome-kcurves-1}

\caption{ASE of the Drosophila connectome graph with quadratic Bezier curves fitted via $K$-curves clustering. The embedding vectors are colored by predicted label.}\label{fig:mbconnectome-kcurves}
\end{figure}
